---
title: "DATA607 - Assignment 5"
author: "Diego Correa"
date: "9/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

1. Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below.
2. Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. 
3. Perform analysis to compare the arrival delays for the two airlines.
4. Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions.


My approach for this assignment is:
  * Gather the data from the CSV file found on GitHub.
  * Clean and transform data
  * Perform analysis of the Airlines and Cities


## Gather Data

```{r include=FALSE}
library(tidyverse, quietly = TRUE)

url <- 'https://github.com/dcorrea614/MSDS/raw/master/Assign5%20-%20flights.csv'
dfFlights <- read.csv(url)
```

## Cleaning the Data

```{r}
dfFlights <- dfFlights %>%
  rename(Airline = ï.., Status = X, 'Los Angeles' = Los.Angeles, 'San Diego' = San.Diego, 
          'San Francisco' = San.Francisco)

colnames(dfFlights)
```
Handling the data

```{r}
# removing the empty row
dfFlights <- dfFlights[-3,]

# Entering missing data inputs
dfFlights[2,1] <- 'Alaska'
dfFlights[4,1] <- 'AM West'

# formatting the data from char to int
dfFlights <- dfFlights %>% 
  mutate(Phoenix = as.integer(str_remove(Phoenix, ',')))

dfFlights <- dfFlights %>% 
  mutate(Seattle = as.integer(str_remove(Seattle, ',')))

dfFlights
```
## Transforming the Data

In order to perform the analysis, I needed to pivot the original data to calculate ratios.


```{r}
# Pivoting the Alaska Airline rows
dfAlaska <- dfFlights[1:2,-1] %>% 
  pivot_longer(!Status, names_to = 'City', values_to = 'count') %>%
  pivot_wider(names_from = Status, values_from = count)

dfAlaska['Airline'] <- 'Alaska'

# Pivoting the AM West Airline rows
dfAMWest <- dfFlights[3:4,-1] %>% 
  pivot_longer(!Status, names_to = 'City', values_to = 'count') %>%
  pivot_wider(names_from = Status, values_from = count)

dfAMWest['Airline'] <- 'AM West'

# Combining the two data frames
df1 <- rbind(dfAlaska, dfAMWest)

df1 <- df1 %>%
  rename(on_time = 'on time')

df1 <- df1 %>%
  select(Airline, City, on_time, delayed)

# Calculating ratios
df1['on_time_ratio'] <- df1['on_time'] / (df1['delayed'] + df1['on_time'])
df1['delayed_ratio'] <- df1['delayed'] / (df1['delayed'] + df1['on_time'])

df1 <- df1 %>%
  select(Airline, City, on_time, delayed, on_time_ratio, delayed_ratio)

df1
```

## Performing Analyses of Total Amounts



```{r}
ggplot(data = df1) + 
  geom_bar(mapping = aes(x = Airline, y = on_time, fill = 'on_time'), stat = 'identity') +
  geom_bar(mapping = aes(x = Airline, y = delayed, fill = 'delayed'), stat = 'identity') + 
  ylab('counts')

ggplot(data = df1) + 
  geom_bar(mapping = aes(x = City, y = on_time, fill = Airline), stat = 'identity', position = 'dodge') +
  coord_flip()
```

## Performing Analysis on Ratios

```{r}
ggplot(data = df1) + 
  geom_bar(mapping = aes(x = City, y = delayed_ratio, fill = Airline), stat = 'identity', position = 'dodge') + 
  coord_flip()

ggplot(data = df1) + 
  geom_bar(mapping = aes(x = City, y = on_time_ratio, fill = Airline), stat = 'identity', position = 'dodge') + 
  coord_flip()
```

### Total Delayed vs Delayed Ratio

```{r}
ggplot(data = df1, mapping = aes(x = delayed, y = delayed_ratio, color = Airline, label = City)) + 
  geom_point() +
  geom_text() +
  geom_smooth(method=lm, aes(fill = Airline))
```

## Conclusion

This last graph illustrates that for areas AM West has higher total delayed flights, they have a lower delayed ratio. The inverse is true for Alaska, where the higher total delayed flights have the higher delayed ratio.

