---
title: "DATA607 - Assingment9"
author: "Diego Correa"
date: "10/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis
You’ll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

I started by creating an account using the above link.  Next, I set up a key and familiarized myself with the API documentation.  Once I was able to access the API through R, I separated the data, transformed, then combined it back together.

##Libraries

```{r include=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
```


## Accessing API

```{r}
url <- 'https://api.nytimes.com/svc/movies/v2/reviews/all.json'
key <- 'GAC1iHZlrjry6V65NHG3DGtCkCOCiikh'

movie_review_data <- GET(url, query = list('api-key' = key))
movie_review_content <- content(movie_review_data, 'text')

movie_review_content <- fromJSON(movie_review_content)

# names of the arrays from the API call
(c_names <- names(movie_review_content))
```
## Viewing the Data Stored

We see that within the JSON file, there are five arrays of data.  Let's see what is in each one of the arrays and transform the data that we want into a data frame

```{r}
df1 <- movie_review_content[c_names[1]]
str(df1)

df2 <- movie_review_content[c_names[2]]
str(df2)

df3 <- movie_review_content[c_names[3]]
str(df3)

df4 <- movie_review_content[c_names[4]]
str(df4)

df5 <- movie_review_content[c_names[5]]
str(df5)
```
## Transforming Dataframes

The variable df5 is the winner, however, "link" and "multimedia" are saved as data frames. We need to separate these data frames, transform them, then bind them together. 

```{r}
# we see that the first nine column names have the prefix of "results."
results_df <- as.data.frame(df5)
results_df <- results_df[1:9]

# transforming column names
c_names <- colnames(results_df) %>%
  str_replace('results.', '')

#  replace the column names
(colnames(results_df) <- c_names)

link_df <- df5$results$link %>%
  as.data.frame()

multimedia_df <- df5$results$multimedia %>%
  as.data.frame()
```
## Combining the Data into a Data Frame

Finally, we can combine our results together into one data frame.

```{r}
movie_df <- cbind(results_df, link_df, multimedia_df)
str(movie_df)
```

## Conclusion

We see that the data stored in the link and mulitmedia data frames are redudant and not needed.
The information found in the results data frame holds information that is easier to read and understand.
