---
title: "DATA607 - Project 1"
author: "Diego Correa"
date: "8/28/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Description

In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605


## Loading and Reading the Data
Firstly, I download the lirbaries needed to manipulate the file and data. Next, I go through the data to understand the format.

```{r cars}
library(readr, quietly = TRUE)
library(stringr, quietly = TRUE)

theUrl <- 'https://raw.githubusercontent.com/dcorrea614/MSDS/master/tournamentinfo.txt'

dfTournament <- read_csv(file = theUrl,col_names = FALSE)
head(dfTournament)
tail(dfTournament)
```

## Deconstructing the Data

Here, I notice the pattern of where the data is stored.  The first 4 lines are not part of the dataset.  It is followed by player information and games played, repeating after every 3 lines.  I, then, separate the data into 2 matrices to resemble a more tradition data set.

```{r}
mTournament <- matrix(unlist(dfTournament), byrow=TRUE)

m1 <- mTournament[seq(5,length(mTournament),3)]
head(m1)

m2 <- mTournament[seq(6,length(mTournament),3)]
head(m2)
```

## Capturing the Data

The next step is to capture the feature values in vectors using string manipulation and regular expressions.  All the features can be captured by standard extraction, besides the Average Pre Chess Rating of Opponents.

```{r warning=FALSE}
# matching first numbers
ID <- as.numeric(str_extract(m1, '\\d+'))

# matching the first combination of a letter, any amount of characters and "|"
Name <- str_extract(m1, '[A-z].{1,32}') 

# extracting the name
Name <- str_trim(str_extract(Name, '.+\\s{2,}'))

# matching the first two letters (state) in the second matrix 
State <- str_extract(m2, '[A-Z]{2}') 

# matching at least 1 number, a period, and 1 number
TotalNumberofPoints <- as.numeric(str_extract(m1, '\\d+\\.\\d'))

# matching the combination of "R", any characters and "-"
PreRating <- str_extract(m2, 'R:.{8,}-')

# matching first 4 numbers
PreRating <- as.numeric(str_extract(PreRating, '\\d{1,4}'))

# matching all combinations of 1 letter 2 spaces and any numbers
Rounds <- str_extract_all(m1, '[A-Z]\\s{2,}\\d+')

# matching numbers
Rounds <- str_extract_all(Rounds, '\\d+')
```

## Calculating Average Pre Chess Rating of Opponents

Using two vectors from the previous step, I am able to calculate the Average Pre Chess Rating of Opponents using a loop.

```{r}
AvgOppPreChessRating <- c()

for(i in c(1:length(Rounds))){
  AvgOppPreChessRating[i] <- round(mean(PreRating[as.numeric(Rounds[[i]])]),0)
}
AvgOppPreChessRating

```

## Constructing the Data Frame

```{r}
Project1 <- data.frame(ID,Name,State,TotalNumberofPoints,PreRating,AvgOppPreChessRating)

head(Project1)


```

## Writing the CSV File

```{r}
write_csv(Project1, 'tournament.csv' , append = FALSE)
```

