---
title: "DATA607 - Assignment 3"
author: "Diego Correa"
date: "9/10/2020"
output: html_document
---


## Question 1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset 
[https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], 
provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
#Load readr library and  use to store csv into dataframe
library(readr, quietly = TRUE)
url <- 'https://github.com/fivethirtyeight/data/raw/master/college-majors/majors-list.csv'
dfMajors <- read.csv(file = url)
head(dfMajors)
```
### Filter Major column for "DATA" or "STATISTICS"

```{r}
#Load tidyverse to use the pipe function and stringr to use the str_detect function
library(tidyverse, quietly = TRUE)
library(stringr, quietly = TRUE)

dfMajors %>%
  filter(str_detect(Major, 'DATA|STATISTICS'))
```


## Question 2
Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", 
"cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", 
"mulberry", "olive", "salal berry")

```{r}
#assuming the above data is store as a string
# y <- str_replace_all(x, ' {2,}', ',')
# y <- as.vector(x)
```


## Question 3

Describe, in words, what these expressions will match:

(.)\1\1                                   

"(.)(.)\\2\\1"

(..)\1                                    

"(.).\\1.\\1"

"(.)(.)(.).*\\3\\2\\1"


```{r}
# a. The regular expression will match characters by grouping the first followed
# by the same character twice more
a <- str_extract(sentences, '(.)\\1\\1') %>% na.omit()
a[1]

# b. The regular expression will match characters by grouping two characters followed 
# by the same two characters in reverse 
b <- str_extract(sentences, '(.)(.)\\2\\1') %>% na.omit()
b[1]

# c. The regular expression will match characters by two characters togerther followed 
# by the same two characters in the same order
c <- str_extract(sentences, '(..)\\1') %>% na.omit()
c[1]

# d. The regular expression will match characters by grouping the first character,
# followed by any character, followed by the first character, followed by any character
# followed by the first character
d <- str_extract(sentences, '(.).\\1.\\1') %>% na.omit()
d[1]

# e. The regular expression will group three characters followed by a character 0 or
# more times followed by the same first three characters in reverse order
e <- str_extract(sentences, '(.)(.)(.).*\\3\\2\\1') %>% na.omit()
e[1]
```

### Question 4

Construct regular expressions to match words that:

Start and end with the same character.

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)


```{r}
words <- c('church', 'statstics', 'eleven')
# '^(.).*\\1$'
aa <- str_match_all(words, '^(.).*\\1$')
aa

# '(.)(.).*\\1\\2'
bb <- str_match_all(words, '(.)(.).*\\1\\2')
bb

# '(.).*\\1.*\\1'
cc <- str_match_all(words, '(.).*\\1.*\\1')
cc
```



