---
title: "DATA607-Assignment2"
author: "Diego Correa"
date: "9/3/2020"
output:
  html_document: default
  pdf_document: default
---

# Introduction

Choose six recent popular movies. Ask at least five people that you know (friends, 
family, classmates, imaginary friends if necessary) to rate each of these movies 
that they have seen on a scale of 1 to 5. Take the results (observations) and 
store them in a SQL database of your choosing. Load the information from the SQL
database into an R dataframe.

Pior to created code, I created 2 CSV files and uploaded them into GitHub.  
Additionally, I created and connected to a cloud-based database.

In the below code, I first captured the data on CSV files into dataframes. 
Followed by establishing a connection to a dattabase. I, then, queried the tables 
and saved the results to a dataframe.  Lastly, I performed analyses on the results.


## Connecting to Data

Here, I loaded the readr package to read the csv files on GitHub and saved them 
to dataframes.

```{r}
library(readr)

friendsUrl <- 'https://raw.githubusercontent.com/dcorrea614/MSDS/master/assignment2%20-%20friends.csv'
ratingURL <- 'https://raw.githubusercontent.com/dcorrea614/MSDS/master/assignment2%20-%20rating.csv'

dfFriends <- read_csv(file = friendsUrl)
dfRating <- read_csv(file = ratingURL)

dfFriends
dfRating
```

## Connecting to Database

To connect to the cloud based database, I used the RMySQL package.

```{r}
library(RMySQL)
con <- dbConnect(MySQL(),
                 user = 'root',
                 host = '34.122.92.218',
                 dbname = 'DATA607')

summary(con)
```

## Creating Tables

Dumping the dataframes into tables.

```{r}
dbWriteTable(con, 'friends', dfFriends, overwrite = TRUE)
dbWriteTable(con, 'rating', dfRating, overwrite = TRUE)
```

## Query Table

Query the tables and joined on FriendsID.

```{r}
res <- dbGetQuery(con, 'select concat(FirstName, " ", LastName) as Friend, Movie, Rating 
                  from rating inner join friends on rating.FriendsID = friends.FriendsID;')
res
```

## Analyze Results

Using the psych package, we are able to see summary statics of the Rating, grouped by Movie.

Finally, I used the ggplot package to visually see the ratings by Movie.

```{r}
library(psych)
describeBy(res$Rating, group = res$Movie)

library(ggplot2)
ggplot(res, aes(x = Rating)) + geom_histogram(fill = 'blue', binwidth = 1) + 
  facet_wrap(~Movie)

```

## Conclusion

Based on the info, Jojo Rabbit and Joker have the highest average ratings, while 
The Irishman received the worst.
